Inference of Bifurcations with Differentiable Continuation

### 2021-02-24 docs/article.tex
Updates from Overleaf

In order for predicted bifurcations $\predictions(\theta)$ to match targets $\targets$ we need to evaluate some form of error term on the bifurcation curve defined by $\rates(z)=0$. We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions can be identified by look for points $p$ along the curve where the determinant $|\frac{\partial\rates}{\partial u}|=0$. We can constrain the cost function to be evaluated in the vicinity of predicted bifurcation points by weighting the likelihood by a function $\omega_\theta(z|\alpha)$ sharply peaked where $|\frac{\partial\rates}{\partial u}|=0$ and going to zero elsewhere. Similarly, we can construct a mixture distribution $\mathcal{N}(p|\targets,\varepsilon)$ peaking at the locations of targets $\targets$ and falling to zero elsewhere. If we choose Gaussian basis functions for both the target mixture and prediction weighting, the likelihood $L_{\theta}(z|\targets)$ takes the form
\begin{align}
    L_{\theta}(z|\targets) := 
    \frac{\omega_\theta(z|\alpha)\,\mathcal{N}(p|\targets,\varepsilon)}{\int_{\rates(z)=0} \omega_\theta(z|\alpha)\,\mathrm{d}z}\label{eq:likelihood}\qquad\qquad\qquad\\
    \mathcal{N}(p|\targets,\varepsilon):=\frac{\sum_{p'\in\targets}\mathbb{e}^{-|p'-p|^2/\varepsilon}}{|\targets|\sqrt{\varepsilon\pi}}
    \quad
    \omega_\theta(z|\alpha):= \mathbb{e}^{-\left|\frac{\partial F_{\theta}}{\partial u}\right|^2/\alpha}
\end{align}
where $\varepsilon$ and $\alpha$ are precision hyper-parameters. Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ along the curve is indeed maximal when the bifurcations $\predictions(\theta)$ of the saddle-node model match the targets $\targets$.

@@

>>>>

In order for predicted bifurcations $\predictions(\theta)$ to match targets $\targets$ we need to evaluate some form of error term on the bifurcation curve defined by $\rates(z)=0$. We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions can be identified by look for points $p$ along the curve where the determinant $|\frac{\partial\rates}{\partial u}|=0$. We can constrain the cost function to be evaluated in the vicinity of predicted bifurcation points by weighting the likelihood by a function $\omega_\theta(z|\alpha)$ sharply peaked where $|\frac{\partial\rates}{\partial u}|=0$ and going to zero elsewhere. Similarly, we can construct a mixture distribution $\mathcal{N}(p|\targets,\varepsilon)$ peaking at the locations of targets $\targets$ and falling to zero elsewhere. If we choose Gaussian basis functions for both the target mixture and prediction weighting, the likelihood $L_{\theta}(z|\targets)$ takes the form
\begin{align}
    L_{\theta}(z|\targets) := 
    \lim_{\alpha\rightarrow\infty}\frac{\omega_\theta(z|\alpha)\,\mathcal{N}(p|\targets,\varepsilon)}{\int_{\rates(z)=0} \omega_\theta(z|\alpha)\,\mathrm{d}z}\label{eq:likelihood}\qquad\qquad\qquad\\
    \mathcal{N}(p|\targets,\varepsilon):=\frac{\sum_{p'\in\targets}\mathbb{e}^{-|p'-p|^2/\varepsilon}}{|\targets|\sqrt{\varepsilon\pi}}
    \quad
    \omega_\theta(z|\alpha):= \mathbb{e}^{-\alpha\left|\frac{\partial F_{\theta}}{\partial u}\right|^2}
\end{align}
where $\varepsilon$ and $\alpha$ are precision hyper-parameters. Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ along the curve is indeed maximal when the bifurcations $\predictions(\theta)$ of the saddle-node model match the targets $\targets$.

Alpha justified as being large by balance between numerical instability due to normalisation and desire to constrain the loss to regions where determinant is zero.

@@

\begin{figure}[H]
\centering{}
\captionsetup{justification=centering}
\includegraphics[width=6cm]{saddle-landscape.png}
\includegraphics[width=6cm]{saddle-optima.png}
\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ optimised with respect to targets $\targets=\{-1,1\}$ in observation region $\Omega\in[-2,2]$. Left panel: trajectories in white approaching black line of global optima.}
\label{fig:saddle-node:results}
\end{figure}

\subsection{Hyperparameters}
Discuss $\alpha$, $\varepsilon$ and $\lambda$.
